% ===========================================================
% Meta-Architecture Specification
% ===========================================================

\section*{Meta-Architecture Specification}

The meta-architecture describes how the components of EFC interact as
a coherent thermodynamic system.  
The architecture ties together the entropy field $S(\mathbf{x})$,
the energy-flow potential $E_f$, the effective speed of light $c(S)$,
and the expansion rate $H(E_f,S)$.

% -----------------------------------------------------------
\subsection*{Core Dependency Graph}

At the structural level, EFC is defined by the following chain:

\[
S(\mathbf{x})
\;\Longrightarrow\;
E_f(\mathbf{x})
\;\Longrightarrow\;
\Phi(E_f,S)
\;\Longrightarrow\;
c(S)
\;\Longrightarrow\;
H(E_f,S)
\;\Longrightarrow\;
\text{Observables}.
\]

Each transformation preserves thermodynamic consistency and constrains
subsequent layers.

% -----------------------------------------------------------
\subsection*{Thermodynamic Integration}

The architecture integrates three thermodynamic components:

\begin{enumerate}
    \item \textbf{Local structure:}  
    entropy gradients and equilibrium zones define haloes.

    \item \textbf{Transport properties:}  
    $c(S)$ determines photon propagation and lensing.

    \item \textbf{Global evolution:}  
    the expansion rate $H$ follows directly from $E_f$ and $S$.
\end{enumerate}

This eliminates the need for dark matter and dark energy as separate,
non-thermodynamic entities.

% -----------------------------------------------------------
\subsection*{Stability Conditions}

A key architectural feature is the set of stability conditions:

\[
\nabla E_f \approx 0 \quad \Rightarrow \quad \text{quasi-stable halo},
\]

\[
|\nabla S| \text{ small} \quad \Rightarrow \quad \text{structural persistence}.
\]

These conditions arise naturally from the entropy field and do not
require additional free parameters.

% -----------------------------------------------------------
\subsection*{Architectural Summary}

The meta-architecture enforces:

\begin{itemize}
    \item coherent interaction between thermodynamic subsystems,
    \item strict mapping across all cosmological observables,
    \item predictive constraints from $S$ and $E_f$ alone,
    \item model reduction without surplus entities.
\end{itemize}

This defines EFC as an integrated, thermodynamic cosmology.
