% =====================================================================
% Energy-Flow Cosmology — C0 Information Model (Master Specification)
% =====================================================================

\section*{EFC–C$_0$ Information Model: Formal Specification}

\subsection*{1. Purpose}

The C$_0$ model defines the informational resonance structure in
Energy-Flow Cosmology (EFC). It specifies how local information capacity,
flow, and resonance emerge from the interaction between the entropy
field $S$, the energy–flow field $E_f$, and grid resistance $R$.

\subsection*{2. Core Fields}

The model uses:

\begin{itemize}
    \item Entropy field: $S(\mathbf{x}, t)$
    \item Energy–flow field: $E_f(\mathbf{x}, t)$
    \item Grid resistance: $R(\mathbf{x}, t)$
    \item Information capacity: $C(\mathbf{x}, t)$
    \item Information resonance scalar: $I_0(\mathbf{x}, t)$
\end{itemize}

\subsection*{3. Information Capacity}

Information capacity decreases monotonically with entropy:

\[
C(\mathbf{x}, t) = C_0 \,(1 - S(\mathbf{x}, t))
\]

Low entropy yields high representational bandwidth.

\subsection*{4. Information Flux}

Informational energy flux is defined as:

\[
\Phi_I = E_f \cdot \nabla S.
\]

Regions with strong energy–flow aligned with entropy gradients carry
the highest informational flux.

\subsection*{5. Information Resonance}

The fundamental C$_0$ resonance quantity is:

\[
I_0 = \gamma \, |E_f \cdot \nabla S|.
\]

Interpretation:
\begin{itemize}
    \item $I_0$ measures local information-processing intensity.
    \item Resonance peaks where energy flow opposes or sharpens entropy gradients.
    \item $I_0$ is the foundational measure used for consciousness and cognition mappings.
\end{itemize}

\subsection*{6. Grid Resistance Coupling}

Resistance modulates informational propagation:

\[
R = R(S, \rho),
\]

and the effective propagation velocity:

\[
c_{\text{eff}} = \frac{1}{1 + R}.
\]

High resistance suppresses information flow.

\subsection*{7. Coupled C$_0$ System}

\[
\begin{aligned}
C &= C_0(1 - S), \\
\Phi_I &= E_f \cdot \nabla S, \\
I_0 &= \gamma \, |E_f \cdot \nabla S|, \\
c_{\text{eff}} &= (1 + R)^{-1}.
\end{aligned}
\]

This defines the full informational layer of EFC.

\subsection*{8. Role in the EFC Framework}

C$_0$ connects:
\begin{itemize}
    \item local entropy structure (EFC–S)
    \item dynamical evolution (EFC–D)
    \item structural emergence (EFC–H)
    \item cognitive and representational bandwidth
    \item information propagation across the grid
\end{itemize}

This module functions as the computational core for all EFC information-layer analysis.

